---
title: "Simple document"
author: "Helen Peng" 
output: word_document
  # html_document:
  #   toc: true
  #   toc_float: true
---

I'm an R Markdown document!

# Section 0: libraries

```{r load_libaries, echo = F, message = F}
library(tidyverse)
```

# Section 1

Here's a **code chunk** that samples from a *normal distribution*:

```{r}
samp = rnorm(100)
length(samp)
```

# Section 2

I can take the mean of the sample, too! The mean is `r mean(samp)`.

# Section 3: a tibble

```{r}
plot_df = tibble(
  x = rnorm(1000, sd = .5),
  y = 1 + 2 * x + rnorm(1000)
)

head(plot_df)
```

# Section 4: plots

These are plots from are random sample.

```{r echo = FALSE}
ggplot(plot_df, aes(x = x)) + geom_histogram()
ggplot(plot_df, aes(x = x, y = y)) + geom_point()
```

# Section 5: Learning Assessment 2

Write a named code chunk that creates a dataframe comprised of: a numeric variable containing a random sample of size 500 from a normal variable with mean 1; a logical vector indicating whether each sampled value is greater than zero; and a numeric vector containing the absolute value of each element. Then, produce a histogram of the absolute value variable just created. Add an inline summary giving the median value rounded to two decimal places. What happens if you set eval = FALSE to the code chunk? What about echo = FALSE?

This plot shows the distribution of the absolute value of $X \sim N(1,1)$ 

```{r}
set.seed(0)
assessment_df = tibble(
  num_var = rnorm(500, mean = 1),
  log_var = num_var > 0,
  abs_var = abs(num_var),
)

ggplot(assessment_df, aes(x = abs_var)) +
  geom_histogram()

median_sample = median(assessment_df$num_var) # my solution (don't use this because using the $ there is a possibility that it goes back and edit the df so it's better to just use the function pull)
median_samp = median(pull(assessment_df, num_var)) # actual solution (tidyverse solution)

median_sample == median_samp
median_sample
```

The median value is `r round(median_sample, digits = 2)`

why is my in-line different from the code chunk?? (probably because i didn't set seed? yeah that's why)

Calculating the median value inline `r round(median(pull(assessment_df,num_var)), digits = 2)`

# Section 6: formatting

## Text formatting

*italic* or *italic* **bold** or **bold** `code` superscript^2^ and subscript~2~

superscript^3^ and subscript~3~

## Headings

# 1st Level Header

## 2nd Level Header

### 3rd Level Header

## Lists

- Bulleted list item 1

- Item 2

  - Item 2a

  - Item 2b

1.  Numbered list item 1

2.  Item 2. The numbers are incremented automatically in the output.

## Tables

| First Header | Second Header |
|--------------|---------------|
| Content Cell | Content Cell  |
| Content Cell | Content Cell  |

# Section 7: Learning Assessment 3

After the previous code chunk, write a bullet list given the mean, median, and standard deviation of the original random sample.

* The median value is `r round(median(pull(assessment_df,num_var)), digits = 2)`

* The mean value is `r round(mean(pull(assessment_df,num_var)), digits = 2)`

* The standard deviation is `r round(sd(pull(assessment_df,num_var)), digits = 2)`
