---
title: "02_data_manipulation"
output: github_document
---

This file is for doing data manipulation.

```{r}
library(tidyverse)
```

Load the litters and pups datasets.

```{r}
litters_df = 
  read_csv(file = "data/FAS_litters.csv",
           na = c("","NA","."))

litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv(file = "data/FAS_pups.csv",
           na = c("","NA","."),
           skip = 3)

pups_df = janitor::clean_names(pups_df)

# most people say <- is better  because it only does variable assignment whereas = is different if it's inside a function mean(x = ...) but tbh it doesn't really matter

```


## `select` some variables
```{r}
select(litters_df, group, litter_number)

select(litters_df, group:gd_of_birth)

select(litters_df, -group)

select(litters_df, starts_with("gd"))

select(litters_df, starts_with("gd"), group)

select(litters_df, GROUP = group, everything())

select(litters_df, -starts_with("gd"))
```

a couple of alternatives
```{r}
rename(litters_df, GROUP = group)

relocate(litters_df, litter_number, gd0_weight) # move variables to the front
```

In the pups data select litter number, sex, and PD ears.
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter` some rows
```{r}
filter(litters_df, gd_of_birth == 19)

filter(litters_df, pups_born_alive <= 5)

filter(litters_df, pups_born_alive != 6)

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, group %in% c("Con7", "Con8"), pups_born_alive < 7)

filter(litters_df, pups_born_alive < 4 | pups_born_alive > 8)

drop_na(litters_df)

drop_na(litters_df, gd0_weight)
```

In the pups dataset, keep only rows where PD walk is less than 11 and sex is 2.
```{r}
filter(pups_df, pd_walk < 11, sex == 2)
```

## `mutate` to change some columns

let's add and change some columns
```{r}
mutate(litters_df, 
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain ^ 2,
       group = str_to_lower(group))
```

In the pups dataset, create a variable that subtracks 7 from pd walk.

```{r}
mutate(pups_df,
       pd_walk_minus_7 = pd_walk - 7)
```

## `arrange` the data

```{r}
arrange(litters_df, pups_born_alive)

arrange(litters_df, desc(pups_born_alive))
```

## Do multiple steps...

```{r}
# this is bad
litters_df = read_csv(file = "data/FAS_litters.csv", na = c("","NA","."))
litters_rename_df = janitor::clean_names(litters_df)
litters_with_gd_df = select(litters_rename_df, group, starts_with("gd"))
litters_no_na_df = drop_na(litters_with_gd_df)
litters_wt_gain_df = mutate(litters_no_na_df, wt_gain = gd18_weight - gd0_weight)

## this is even worse
select(janitor::clean_names(read_csv(file = "data/FAS_litters.csv", na = c("","NA","."))), group, starts_with("gd"))

## this is good!!
litters_df = 
  read_csv(file = "data/FAS_litters.csv", na = c("","NA",".")) |> 
  janitor::clean_names() |> 
  select(group, starts_with("gd")) |> 
  drop_na() |> 
  mutate(
    wt_gain =  gd18_weight - gd0_weight,
    group = str_to_lower(group)
  )

```

Load pups, clean names, drop missing, keep litter number and pd variables, and add pd walk minus 7.
```{r}
pups_df = 
  read_csv(file = "data/FAS_pups.csv", na = c("","NA","."), skip = 3) |> 
  janitor::clean_names() |> 
  drop_na() |> 
  select(litter_number, starts_with("pd")) |> 
  mutate(
    pd_walk_minus_7 = pd_walk - 7
  )
```

sometime the first thing isn't a dataframe

```{r}
litters_df |> 
  filter(group %in% c("con7", "con8")) |> 
  lm(gd18_weight ~ gd0_weight, data = _)
```

`select` vs. `pull`
```{r}
litters_df |> 
  select(group) 

litters_df |> 
  pull(group)

# select always give a df but pull gives a vector
# if you have a df you should never take something out of the dataframe so dont use pull

litters_df$group # don't do this, never do this

# difference between pull and $ functionally not much, subtle differences not important. But it is better to use pull because it's clear you are taking something out of a dataframe. 

```

pipes and view
```{r}
# litters_df |> 
#   filter(group %in% c("con7", "con8")) |> 
#   view()

# you can view dataset but remember remove it or just do it in the console
```

