---
title: "02_data_manipulation"
output: github_document
---

This file is for doing data manipulation

```{r}
library(tidyverse)
```

Load the litters and pups datasets.

```{r}
litters_df = 
  read_csv(file = "data/FAS_litters.csv",
           na = c("","NA","."))

litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv(file = "data/FAS_pups.csv",
           na = c("","NA","."),
           skip = 3)

pups_df = janitor::clean_names(pups_df)

# most people say <- is better  because it only does variable assignment whereas = is different if it's inside a function mean(x = ...) but tbh it doesn't really matter

```


## `select` some variables
```{r}
select(litters_df, group, litter_number)

select(litters_df, group:gd_of_birth)

select(litters_df, -group)

select(litters_df, starts_with("gd"))

select(litters_df, starts_with("gd"), group)

select(litters_df, GROUP = group, everything())

select(litters_df, -starts_with("gd"))
```

a couple of alternatives
```{r}
rename(litters_df, GROUP = group)

relocate(litters_df, litter_number, gd0_weight) # move variables to the front
```

In the pups data select litter number, sex, and PD ears.
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter` some rows
```{r}
filter(litters_df, gd_of_birth == 19)

filter(litters_df, pups_born_alive <= 5)

filter(litters_df, pups_born_alive != 6)

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, group %in% c("Con7", "Con8"), pups_born_alive < 7)

filter(litters_df, pups_born_alive < 4 | pups_born_alive > 8)

drop_na(litters_df)

drop_na(litters_df, gd0_weight)
```

In the pups dataset, keep only rows where PD walk is less than 11 and sex is 2.
```{r}
filter(pups_df, pd_walk < 11, sex == 2)
```

## `mutate` to change some columns

let's add and change some columns
```{r}
mutate(litters_df, 
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain ^ 2,
       group = str_to_lower(group))
```

In the pups dataset, create a variable that subtracks 7 from pd walk.

```{r}
mutate(pups_df,
       pd_walk_minus_7 = pd_walk - 7)
```

## `arrange` the data

```{r}
arrange(litters_df, pups_born_alive)

arrange(litters_df, desc(pups_born_alive))
```

